---
title: "RShiny Exercise 1 - Solutions"
author: "Jacques van Helden"
date: '`r Sys.Date()`'
output:
  html_document:
    self_contained: no
    fig_caption: yes
    highlight: zenburn
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float: yes
    code_folding: "show"
  beamer_presentation:
    colortheme: dolphin
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    fonttheme: structurebold
    highlight: tango
    incremental: no
    keep_tex: no
    slide_level: 2
    theme: Montpellier
    toc: yes
  revealjs::revealjs_presentation:
    theme: night
    transition: none
    self_contained: true
    css: ../slides.css
  pdf_document:
    fig_caption: yes
    highlight: zenburn
    toc: yes
    toc_depth: 3
  slidy_presentation:
    smart: no
    slide_level: 2
    self_contained: yes
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    incremental: no
    keep_md: yes
    smaller: yes
    theme: cerulean
    toc: yes
    widescreen: yes
  ioslides_presentation:
    slide_level: 2
    self_contained: no
    colortheme: dolphin
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    fonttheme: structurebold
    highlight: tango
    smaller: yes
    toc: yes
    widescreen: yes
font-import: http://fonts.googleapis.com/css?family=Risque
subtitle: "Diplôme Interuniversitaire en Bioinformatique intégrative (DU-Bii 2019)"
font-family: Garamond
transition: linear
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## 1. Starting application

Copy, debug and run the first application (app1) presented today.

```{r eval=FALSE, echo=TRUE}
ui <- fluidPage(
  titlePanel("Workshop - Example 1 – Basic  Histogram"),
  sidebarLayout(
    sidebarPanel(
       numericInput(inputId="n",
                    label="Number of observations",
                    value=1000) ),  
       mainPanel(plotOutput("plot"))
  )
)

server <- function(input, output) {
   data <- reactive({
                     x <- rnorm(input$n)
                     x })

   output$plot <- renderPlot({
         	        hist(data(), 50, main="",  xlab="x")
                   })
}
shinyApp(ui = ui, server = server)
```



Solution: [app1.R](app1.R)


## 2. Bin numbers

Change the application app1 to allow the user to enter the numbers of bins when plotting the histogram

Solution: [app2.R](app2.R)

## 3. Server response

Change the application app1 to add some text as a response from server. 

Solution: [app3.R](app3.R)

## 4. Mean and standard deviation

Change app1 so that

- The user can enter different values for the mean and standard deviation. 
- The mean and standard deviation are shown in the mainPanel.
- The user can enter a title for the histogram.

Solution: [app4.R](app4.R)

## 5. Alternative displays

Change app1 so that the user can choose to do a
histogram or a boxplot

Solution: [app5.R](app5.R)


## 6. Do it the way you want it

Just about anything can be made to look the way you want it.

- change the size of the sidebar: (,width=3)
- Change the aspect ratio of the graph:

```{r eval=FALSE}
plotOutput("plot", width = "500px", height = "500px")
```

Solution: [app6.R](app6.R)

## 7. Panel appears and disappears:

An ugly feature of our app: the input field Number of bins only makes sense for the histogram, not for the boxplot, so it should not appear when we do a boxplot.

**Hint:** `conditionalPanel()`


Solution: [app7.R](app7.R)






